

```{r libraries, echo=FALSE}

library(scales)
# library(cowplot)
library(here)
library(vegan)
```


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
```


```{r data-import, echo=FALSE}

source(here::here("R", "hydro_metrics.R"))
source(here::here("R", "hysteresis_calcs.R"))
```


Plot of of cumQ v. totalLoad. Next step here is to get at the residuals.

```{r cumQ-totalLoad}

hydroMetrics %>% 
  inner_join(
    contributingGauges,
    by = c("stormMark")
  ) %>% 
  filter(analyte %in% analytesWithTotal) %>% 
  ggplot(aes(x = cumQ, y = totalLoad)) +
  geom_point(aes(colour = fromLakeMarguerite)) +
  # label storms
  # geom_text(aes(label = stormMark), hjust = "right", vjust = "top", nudge_x = 0) +
  geom_smooth(method = lm, se = FALSE, formula = y ~ x, size = 0.5) +
  # stat_poly_eq(formula = y ~ x, parse = TRUE, size = 3, rr.digits = 3, colour = "black") + 
  # stats with equation
  stat_poly_eq(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~")),
               formula = y ~ x, parse = TRUE, size = 3, rr.digits = 3, colour = "black") + 
  scale_x_log10(labels = scales::scientific) +
  scale_y_log10(labels = scales::scientific) +
  facet_wrap(~ analyte, scales = "free")

```


Plot of Flusing Index v. HImean having joined hydroMetrics + hysteresisIndices +
contributingGauges. Here grouped by `monsoon` but change grouping variable to
whatever parameter is of interest (e.g., `reachLength`).

```{r hysteretic-responses}

hydroMetrics %>% 
  inner_join(
    hysteresisIndices,
    by = c("stormMark", "analyte")
  ) %>% 
  inner_join(
    contributingGauges,
    by = c("stormMark")
  ) %>% 
  ggplot(aes(x = FI, y = HImean)) + 
  lims(x = c(-1, 1), y = c(-1, 1)) +
  geom_vline(xintercept = 0, colour = "gray") +
  geom_hline(yintercept = 0, colour = "gray") +
  geom_point(aes(colour = reachLength)) + # grouping var.
  facet_wrap(~ analyte) +
  theme_minimal()

```


Shotgun scatterplot everything v. everything.

```{r scatterplot}

hydroMetrics %>% 
  inner_join(
    hysteresisIndices,
    by = c("stormMark", "analyte")
  ) %>% 
  inner_join(
    contributingGauges,
    by = c("stormMark")
  ) %>% 
  select(totalLoad:fromInterceptor) %>% 
  pairs()
```


Boxplots of totalLoad per grouping variable, here reachLength but change x to
parameter of interest.

```{r boxplots}

hydroMetrics %>% 
  inner_join(
    hysteresisIndices,
    by = c("stormMark", "analyte")
  ) %>% 
  inner_join(
    contributingGauges,
    by = c("stormMark")
  ) %>% 
  ggplot(aes(x = reachLength, y = totalLoad)) +
  geom_boxplot() +
  facet_wrap(~ analyte, scales = "free")
```


Generate a data frame of hysteresis data for a particular analyte to include in
a ordination.

```{r fn HI and FI per analyte, echo=FALSE}

# Function to generate a data frame of hysteresis data for a particular analyte
# to include in a ordination. Input parameter (chemSpecies) is the quoted name
# of the analyte of interest (e.g., "ClD_LACHAT")

generate_hyst_data <- function(chemSpecies) {
  
  hydroMetrics %>% 
    inner_join(
      hysteresisIndices,
      by = c("stormMark", "analyte")
    ) %>% 
    inner_join(
      contributingGauges,
      by = c("stormMark")
    ) %>% 
    filter(analyte == chemSpecies) %>%
    select(stormMark, HImean, FI) -> chemSpeciesData
  
  chemSpeciesData <- as.data.frame(chemSpeciesData) # to data frame
  rownames(chemSpeciesData) <- chemSpeciesData$stormMark # stormMark to row name
  chemSpeciesData <- chemSpeciesData[,c("HImean", "FI")] # remove stormMark col
  
  return(chemSpeciesData)
  
}

# EXAMPLE
# chemCl <- generate_chem_data("ClD_LACHAT")
# chemCl <- generate_hyst_data("ClD_LACHAT")

```


Generate a data frame of environmental data to include in a ordination.

```{r fn environmental data, echo=FALSE}

# Function to generate a data frame of environmental data. Input parameters are
# (rmcolin), a binary flag indicating if a selection of collinear variables
# should be included (0 (default)) or removed (1), and chemSpecies, the analyte
# of interest. Variables to remove should be adjusted in the function.

generate_env_data <- function(rmcolin, chemSpecies) {
  
  if (missing(rmcolin)) {
    rmcolin <- 0
  }
  
  hydroMetrics %>% 
    inner_join(
      hysteresisIndices,
      by = c("stormMark", "analyte")
    ) %>% 
    inner_join(
      contributingGauges,
      by = c("stormMark")
    ) %>% 
    inner_join(
      ibwPrecipitation %>% select(stormMark, IDW_IDP2),
      by = c("stormMark")
    ) %>% 
    filter(analyte == chemSpecies) %>%
    select(-analyte, -b, -HImean, -CQpeak, -CQinit, -FI) -> envData 
  
  if (rmcolin == 1) {
    
    # from: cor(envData[sapply(envData, is.numeric)], method = "pearson")
    envData <- envData %>% 
      select(-maxC, -maxQ, -cumQ, -IDW_IDP2) 
  }
  
  # format to accommodate vegan
  envData <- as.data.frame(envData) # to data frame
  rownames(envData) <- envData$stormMark # stormMark to row name
  envData <- subset(envData, select = -c(stormMark)) # remove stormMark col
  
  # scale continuous env variables
  varsCont <- envData[sapply(envData, is.numeric)] # isolate continuous vars
  varsCont <- decostand(varsCont, method = 'standardize') # scale continuous vars
  varsCat <- envData[sapply(envData, is.factor)] # isolate categorical vars
  envData <- cbind(varsCont, varsCat) # new envData = scaled continuous + cat
  
  # return 
  return(envData)
  
}

# EXAMPLE
# envData <- generate_env_data(1, "ClD_LACHAT")
```


workflow: NMDS, PCA, and PCA + fitting env. data

* without predictors, RDA is the same as PCA
* in vegan, giving the function rda() dataframe without predictors runs a PCA
very simiar to princomp https://rpubs.com/brouwern/veganpca

```{r ordination per analyte}

analyte <- "ClD_LACHAT"

hystData <- generate_hyst_data(analyte) # generate hysteresis data
envData <- generate_env_data(1, analyte) # generate environmental data

ibwNMDS <- metaMDS(hystData) # NMDS
PCA <- rda(hystData, scale = TRUE) # PCA
PCAenv <- envfit(PCA, envData, permutations = 999) # fit env data to PCA

plot(PCA)
plot(PCAenv)
plot(PCAenv, p.max = 0.05)

```


workflow: generate RDA of hysteresis metrics HImean & FI constrained to
environmental variables.

```{r RDA per analyte}

analyte <- "ClD_LACHAT"
analyte <- "DOC_TOC"
analyte <- "NO3D_LACHAT"
analyte <- "PO4D_LACHAT"
analyte <- "NH4_LACHAT"
analyte <- "NaD_ICP"

hystData <- generate_hyst_data(analyte) # generate hysteresis data
envData <- generate_env_data(rmcolin = 1,
                             chemSpecies = analyte) # generate environmental data
# choose one:
  envData <- envData[sapply(envData, is.numeric)] # continuous only
  envData <- subset(envData, select = -reachLength) # remove reachLength factor

# full RDA
fullRDA <- rda(hystData ~ ., envData, scale = TRUE)
anova(fullRDA)
anova(fullRDA, by = "margin")
anova(fullRDA, by = "axis")
summary(fullRDA)[['cont']]
plot(fullRDA)
vif.cca(fullRDA)

envfit(rda(hystData, scale = TRUE), envData) # check significance of individual vars

# forward-backward selection
m0 <- rda(hystData ~ 1, envData, scale = TRUE)
finalRDA <- ordistep(m0, scope = formula(fullRDA))
vif.cca(finalRDA)
anova(finalRDA)
anova(finalRDA, by = "margin")
anova(finalRDA, by = "axis")
plot(finalRDA)
summary(finalRDA)[['cont']]

# variance partitioning

```


generally:
Better fits to reactive species, and particularly with continuous vars only, in fact the nitrate model is significant. Storm duration has influence. 

chloride
* overall p = 0.3
* optimal model includes only reachLength

sodium
* overall 0.52
* optimal model includes only reachLength

nitrate:
* overall 0.079; 0.037 with only continuous
* storm duration signifant when including only continuous
* optimal model includes only stormDuration, but antecedentDay was significant among margins (low variance for both, though)

phophate
* overall 0.334
* final model includes only fromInterceptor

ammonium
* overall 0.061
* final model includes only reachLength

DOC
* overall 0.062
* optimal model includes only reachLength

```{r RDA on each FI/HI }

analytesSansTotals <- c("ClD_LACHAT",
                        "DOC_TOC",
                        "NO3D_LACHAT",
                        "PO4D_LACHAT",
                        "NH4_LACHAT",
                        "NaD_ICP"
                        # "NO3T", # missing values
                        # "PO4T" # missing values
)

speciesDataFI <- hydroMetrics %>% 
  inner_join(
    hysteresisIndices,
    by = c("stormMark", "analyte")
  ) %>% 
  inner_join(
    contributingGauges,
    by = c("stormMark")
  ) %>% 
  select(analyte, stormMark, FI) %>%
  filter(analyte %in% analytes) %>% 
  spread(key = analyte, value = FI) %>%
  rename(rowname = stormMark) %>% 
  column_to_rownames()

fullRDA <- rda(speciesDataFI ~ ., envData, scale = TRUE)
plot(fullRDA)
# PCA <- rda(speciesDataFI, scale = TRUE) # PCA

speciesDataHI <- hydroMetrics %>% 
  inner_join(
    hysteresisIndices,
    by = c("stormMark", "analyte")
  ) %>% 
  inner_join(
    contributingGauges,
    by = c("stormMark")
  ) %>% 
  select(analyte, stormMark, HImean) %>%
  filter(analyte %in% analytes) %>% 
  spread(key = analyte, value = HImean) %>%
  rename(rowname = stormMark) %>% 
  column_to_rownames()

fullRDA <- rda(speciesDataHI ~ ., envData, scale = TRUE)
PCA <- rda(speciesDataHI, scale = TRUE) # PCA

# all FI/HI require envData that is not specific to any single analyte
envData <- hydroMetrics %>% 
  select(stormMark, stormDuration, cumQ, maxQ, monsoon, antecedentDay) %>% 
  mutate(monsoon = as.character(as.numeric(monsoon))) %>% 
  group_by(stormMark) %>% 
  summarise_all(max) %>% 
  mutate(monsoon = as.factor(monsoon)) %>% 
  inner_join(
    contributingGauges,
    by = c("stormMark")
  ) %>% 
  inner_join(
    ibwPrecipitation %>% select(stormMark, IDW_IDP2),
    by = c("stormMark")
  )

# format to accommodate vegan
envData <- as.data.frame(envData) # to data frame
rownames(envData) <- envData$stormMark # stormMark to row name
envData <- subset(envData, select = -c(stormMark)) # remove stormMark col

# scale continuous env variables
varsCont <- envData[sapply(envData, is.numeric)] # isolate continuous vars
varsCont <- decostand(varsCont, method = 'standardize') # scale continuous vars
varsCat <- envData[sapply(envData, is.factor)] # isolate categorical vars
envData <- cbind(varsCont, varsCat) # new envData = scaled continuous + cat

# remove reachLength as factor (if desired)
envData <- envDatad %>% select(-reachLength) 

```


```{r}

hydroMetrics %>% 
  filter(analyte == "ClD_LACHAT") %>% 
  lm(cumQ ~ totalLoad, data = .)
  
fit <- lm(cumQ ~ totalLoad, data = hydroMetrics %>% filter(analyte == "ClD_LACHAT")) # Fit the model

```

