---
title: "plotting_hysteresis"
output:
  html_document:
    df_print: paged
date: "2026-03-26"
---

Aims to plot hysteresis per analyte per storm per site

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=12, fig.height=8) 
```


```{r libraries, include=FALSE}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(here)
library(purrr)
```

## Data munging/cleaning

Workflow before this rmd, in order:

* q_storms.R: import discharge from Lake Margherite, Silverado, & Curry, delineate unique storms, and interpolate Q
* merge_Q_chems.R: merge discharge data with chem data per site, produce brief cq plots per analyte per site
* hysteresis_calc_2.R: calculate hysteresis based on python function, removed a lot of storms with not enough data

```{r load data}
curry_hysteresis <- readRDS(here("curry_hysteresis.rds")) # raw product from hysteresis calculation, per analyte per storm
silv_hysteresis <- readRDS(here("silv_hysteresis.rds"))
lakem_hysteresis <- readRDS(here("lakem_hysteresis.rds"))

hysteresis_df <- read.csv(here("hysteresis_summary.csv")) # summearized, combined all sites, includes mean HI per analyte per storm, amongst other things
hi_intervals_df <- read.csv(here("hysteresis_intervals.csv")) # raw HI per interval per site per storm per analyte
```


## Plotting

```{r echo=FALSE, warning=FALSE, message=FALSE}
hi_summary <- hysteresis_df %>%
  group_by(site, analyte) %>%
  summarise(
    median_HI = median(HI_mean, na.rm = TRUE),
    lower_CI  = quantile(HI_mean, 0.25, na.rm = TRUE),  # 25th percentile
    upper_CI  = quantile(HI_mean, 0.75, na.rm = TRUE),  # 75th percentile
    n         = n(),
    .groups = "drop"
  )
hi_summary <- hi_summary %>% filter(!is.na(analyte))
hysteresis_df %>% filter(!is.na(analyte)) %>%  ggplot() +
  geom_jitter(aes(x = site, y = HI_mean),
              width = 0.15, alpha = 0.3, size = 1.5) +
  geom_pointrange(data = hi_summary,
                  aes(x = site, y = median_HI, ymin = lower_CI, ymax = upper_CI),
                  size = 0.8, linewidth = 1.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Median HI by Analyte and Site",
       subtitle = "Points = individual storms, range = IQR",
       x = NULL, y = "HI") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))


#per site
hysteresis_df %>%
  filter(site == "Curry", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Curry - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

hysteresis_df %>%
  filter(site == "Silverado", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Silverado - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

hysteresis_df %>%
  filter(site == "Lake Margarita", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Lake Margarita - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

