---
title: "Hysteresis Indicies and Beta of IBW sites: Curry, Silverado, and Lake Marguerite"
output:
  pdf_document: default
  html_document:
    df_print: paged
indent: true
date: "2026-03-26"
---

Aims to plot hysteresis per analyte per storm per site

## Metrics



* *Hysteresis index* is calculated as the median (+/- 95% CI) across 2% intervals of discharge.


    +HI = clockwise (source limitation or nearstream source)


    -HI = counterclockwise (transport limitation or distal source)



* *Beta* is the slope of the CQ relationship on the rising limb of the storm

    CIs on beta are bootstrapped 95% CI
    

**NOTES**


* Lake Marguerite does not have a lot of chems to work with. Do we know if there are more chems than the 624_runoff_chems.csv from the dropbox?

* Similarly, there is a gap of data in Silverado 2014-2018, was there less sampling in that time or is the data elsewhere?

* individual analyte outliers > 3sd above mean concentration per storm were removed for hysteresis and beta calculations, but remain for individual storm plotting at the bottom of the report

* Precipitation data was collected from https://azmet.arizona.edu/azmet/27.htm - precipitation gage at the Desert Ridge station

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=12, fig.height=8) 
```


```{r libraries, include=FALSE}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(here)
library(purrr)
#library(GGally)
library(corrplot)
```

## Data munging/cleaning

Workflow before this rmd, in order:

* q_storms.R: import discharge from Lake Margherite, Silverado, & Curry, delineate unique storms, and interpolate Q
* precipitation.R: import precip data and merge with Q to calculate pre-storm precip metrics
* merge_Q_chems.R: merge discharge data with chem data per site, produce brief cq plots per analyte per site
* hysteresis_calc_2.R: calculate hysteresis based on python function, removed a lot of storms with not enough data. also calculates beta

```{r load data, include=FALSE}

# note - we cannot access google drive API when knitting rmd, so these files will have to be downloaded locally. this can be done by running hysteresis_calc_2.R

curry_hysteresis <- readRDS(here("Data/curry_hysteresis.rds")) # raw product from hysteresis calculation, per analyte per storm
silv_hysteresis <- readRDS(here("Data/silv_hysteresis.rds"))
lakem_hysteresis <- readRDS(here("Data/lakem_hysteresis.rds"))

hysteresis_df <- read.csv(here("Data/hysteresis_summary.csv")) # summearized, combined all sites, includes mean HI per analyte per storm, amongst other things
hi_intervals_df <- read.csv(here("Data/hysteresis_intervals.csv")) # raw HI per interval per site per storm per analyte
beta_df <- read.csv(here("Data/beta_all_sites.csv"))#beta per storm per analyte per site

# from google drive - no API access needed 
curry_cq <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", 
                             "1wWUPyysrln9gyWpfS6vAt2u-MS-LtOyT")) %>% 
  select(-X) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"))
curry_cq[1,1] <- "2008-01-29 00:00:00"

silv_cq <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", 
                             "1J-mW-MfwTfq7sOsy2sn4eejb9cq0ykvR")) %>% 
  select(-X) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"))
silv_cq[1,1] <- "2008-01-29 00:00:00"

lakem_cq <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", 
                            "100HPPcicy6tv7CUrwzDI8Yh4C6FBozhG")) %>% 
  select(-X) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"))
lakem_cq[1,1] <- "2008-01-29 00:00:00"
```

```{r data munging, message=FALSE, warning=FALSE, include=FALSE}
#some quick data munging - need to get storm_date into beta
hysteresis_df <- hysteresis_df %>% mutate(season = ifelse((month(storm_date) >= 10 | month(storm_date) < 5),  "winter", "summer"))

storm_date <- hysteresis_df %>% select(c(site, stormID, storm_date, analyte, season))
beta_df <- left_join(beta_df, storm_date, by = c("site", "stormID", "analyte"))
```


## Plotting Hysteresis

```{r echo=FALSE, warning=FALSE, message=FALSE}
hi_summary <- hysteresis_df %>%
  group_by(site, analyte) %>%
  summarise(
    median_HI = median(HI_mean, na.rm = TRUE),
    lower_CI  = quantile(HI_mean, 0.25, na.rm = TRUE),  # 25th percentile
    upper_CI  = quantile(HI_mean, 0.75, na.rm = TRUE),  # 75th percentile
    n         = n(),
    .groups = "drop"
  )
hi_summary <- hi_summary %>% filter(!is.na(analyte))
# hysteresis_df %>% filter(!is.na(analyte)) %>%  ggplot() +
#   geom_jitter(aes(x = site, y = HI_mean),
#               width = 0.15, alpha = 0.3, size = 1.5) +
#   geom_pointrange(data = hi_summary,
#                   aes(x = site, y = median_HI, ymin = lower_CI, ymax = upper_CI),
#                   size = 0.8, linewidth = 1.2) +
#   geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
#   facet_wrap(~ analyte, scales = "free_y") +
#   labs(title = "Median HI by Analyte and Site",
#        subtitle = "Points = individual storms, range = IQR",
#        x = NULL, y = "HI") +
#   theme_bw() +
#   theme(legend.position = "none",
#         axis.text.x = element_text(angle = 45, hjust = 1))


#per site
hysteresis_df %>%
  filter(site == "Curry", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean, color = season)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Curry - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15, angle = 45, hjust = 1), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 10))

hysteresis_df %>%
  filter(site == "Silverado", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean, color = season)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Silverado - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15, angle = 45, hjust = 1), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 10))

hysteresis_df %>%
  filter(site == "Lake Margarita", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean, color = season)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Lake Marguerite - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15, angle = 45, hjust = 1), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 10))

```

## Plotting Beta


```{r plotting beta, echo=FALSE, warning=FALSE, message=FALSE}

# beta_df %>% filter(!is.na(analyte), analyte != "Pb", outlier==F) %>% ggplot(aes(x = as.Date(storm_date), y = beta, color = site))+
#   geom_point()+
#   geom_errorbar(aes(ymin = boot_ci_lo, ymax = boot_ci_hi),
#                 width = 10, alpha = 0.4) +
#   facet_wrap(~analyte, scales = "free_y") + theme_minimal() +
#   labs(title = "Beta per storm per analyte per site", x = "Storm Date", subtitle = "Bars are 95% confidence intervals")

beta_df %>% filter(site == "Curry", !is.na(analyte), analyte != "Pb", outlier==F) %>% ggplot(aes(x = as.Date(storm_date), y = beta, color = season))+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_errorbar(aes(ymin = boot_ci_lo, ymax = boot_ci_hi),
                width = 10, alpha = 0.4) +
  facet_wrap(~analyte, scales = "free_y") + theme_minimal() +
  labs(title = "Curry Beta per storm per analyte", x = "Storm Date", subtitle = "Bars are 95% confidence intervals")+
  theme(axis.text.x = element_text(size = 15, angle = 45, hjust = 1), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 10))

beta_df %>% filter(site == "Silverado", !is.na(analyte), analyte != "Pb", outlier==F) %>% ggplot(aes(x = as.Date(storm_date), y = beta, color = season))+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_errorbar(aes(ymin = boot_ci_lo, ymax = boot_ci_hi),
                width = 10, alpha = 0.4) +
  facet_wrap(~analyte, scales = "free_y") + theme_minimal() +
  labs(title = "Silverado Beta per storm per analyte", x = "Storm Date", subtitle = "Bars are 95% confidence intervals")+
  theme(axis.text.x = element_text(size = 15, angle = 45, hjust = 1), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 10))

beta_df %>% filter(site == "Lake Margarita", !is.na(analyte), analyte != "Pb", outlier==F) %>% ggplot(aes(x = as.Date(storm_date), y = beta, color = season))+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_errorbar(aes(ymin = boot_ci_lo, ymax = boot_ci_hi),
                width = 10, alpha = 0.4) +
  facet_wrap(~analyte, scales = "free_y") + theme_minimal() +
  labs(title = "Lake Marguerite Beta per storm per analyte", x = "Storm Date", subtitle = "Bars are 95% confidence intervals")+
  theme(axis.text.x = element_text(size = 15, angle = 45, hjust = 1), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 10))

```



```{r, echo=FALSE, warning=FALSE, message=FALSE}


curry_hi_beta <- hysteresis_df %>%
  filter(site == "Curry") %>%
  inner_join(
    beta_df %>% filter(site == "Curry"),
    by = c("stormID", "analyte", "site", "season", "storm_date")
  )

# plot
curry_hi_beta %>%
  filter(!is.na(analyte), !outlier) %>%
  ggplot(aes(x = beta, y = HI_mean, color = season)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd, colour = "red"),
                alpha = 0.4, width = 0.05) +
  geom_errorbar(aes(xmin = boot_ci_lo, xmax = boot_ci_hi, colour = "blue"),
                alpha = 0.4, width = 0.05,
                orientation = "y") +          # <-- replaces geom_errorbarh
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ analyte, scales = "fixed") +
  theme_minimal() + theme(axis.text.x = element_text(size = 15), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15)) +
  labs(title = "Curry - HI vs Beta by Analyte", subtitle = "Each dot is an individual storm",
       x = "Beta",              
       y = "HI")+scale_color_manual(values = c("summer" ="red", "winter" = "blue")) 

silv_hi_beta <- hysteresis_df %>%
  filter(site == "Silverado") %>%
  inner_join(
    beta_df %>% filter(site == "Silverado"),
    by = c("stormID", "analyte", "site", "season", "storm_date")
  )

# plot
silv_hi_beta %>%
  filter(!is.na(analyte), !outlier) %>%
  ggplot(aes(x = beta, y = HI_mean, color = season)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd, colour = "red"),
                alpha = 0.4, width = 0.05) +
  geom_errorbar(aes(xmin = boot_ci_lo, xmax = boot_ci_hi, colour = "blue"),
                alpha = 0.4, width = 0.05,
                orientation = "y") +          # <-- replaces geom_errorbarh
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ analyte, scales = "fixed") +
  theme_minimal() + theme(axis.text.x = element_text(size = 15), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15)) +
  labs(title = "Silverado - HI vs Beta by Analyte", subtitle = "Each dot is an individual storm",
       x = "Beta",              
       y = "HI")+scale_color_manual(values = c("summer" ="red", "winter" = "blue")) 

lakem_hi_beta <- hysteresis_df %>%
  filter(site == "Lake Margarita") %>%
  inner_join(
    beta_df %>% filter(site == "Lake Margarita"),
    by = c("stormID", "analyte", "site", "season", "storm_date")
  )

# plot
lakem_hi_beta %>%
  filter(!is.na(analyte), !outlier) %>%
  ggplot(aes(x = beta, y = HI_mean, color = season)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd, colour = "red"),
                alpha = 0.4, width = 0.05) +
  geom_errorbar(aes(xmin = boot_ci_lo, xmax = boot_ci_hi, colour = "blue"),
                alpha = 0.4, width = 0.05,
                orientation = "y") +          # <-- replaces geom_errorbarh
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ analyte, scales = "fixed") +
  theme_minimal() + theme(axis.text.x = element_text(size = 15), strip.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15)) +
  labs(title = "Lake Margarita - HI vs Beta by Analyte", subtitle = "Each dot is an individual storm",
       x = "Beta",              
       y = "HI")+scale_color_manual(values = c("summer" ="red", "winter" = "blue")) 
```


## Plotting CQ
```{r cq plots, echo=FALSE, message=FALSE, warning=FALSE}
## attempting to make function for this - from sycburns
plot_c_q_norm <- function(cq_data, analyte_col) {
  curr_site <- cq_data %>% 
    filter(analyte == analyte_col) %>% 
    filter( !is.na(cfs)) 
  
  if (nrow(curr_site) == 0) {
    message(paste("Skipping", site, "- no valid data"))
    return(NULL)
  }
  
  if (!"Time_hours" %in% names(curr_site)) {
    curr_site <- curr_site %>%
     # group_by(stormID) %>% 
      mutate(Time_hours = as.numeric(difftime(datetime, min(datetime), units = "hours")))
  }
  # # Remove top 10% of cfs values per Event_Date (storm)
  # curr_site <- curr_site %>%
  #   group_by(Event_Date) %>%
  #   filter(cfs <= quantile(cfs, 0.9, na.rm = TRUE)) %>%
  #   ungroup()
  
  #Normalize within each storm
  curr_site <- curr_site %>%
    group_by(stormID) %>%
    mutate(
      Level_norm = (cfs - min(cfs, na.rm = TRUE)) /
        (max(cfs, na.rm = TRUE) - min(cfs, na.rm = TRUE)),
      Chem_norm = (.data$mean_conc - min(.data$mean_conc, na.rm = TRUE)) /
        (max(.data$mean_conc, na.rm = TRUE) - min(.data$mean_conc, na.rm = TRUE)),
      Time_norm = (Time_hours - min(Time_hours, na.rm = TRUE)) /
        (max(Time_hours, na.rm = TRUE) - min(Time_hours, na.rm = TRUE))
    ) %>%
    ungroup() %>%
    filter(!is.na(.data$mean_conc), !is.na(cfs)) 
  
  #Plot normalized cq relationship per storm
  p <- curr_site %>% filter(!is.na(curr_site$Level_norm), !is.na(curr_site$Chem_norm)) %>% ggplot(aes(x = Level_norm, y = Chem_norm, shape = outlier_analyte)) + 
    geom_point(aes(color = Time_norm)) + 
    facet_wrap(~stormID, scales = "free") + 
    scale_color_gradientn(colours = rainbow(7)) + 
    labs( title = paste(curr_site$Site[1], "-", analyte_col), 
          x = "Normalized Stage (0–1)", 
          y = paste("Normalized", analyte_col, "(0–1)"), 
          color = "Time (within storm)" ) + 
    theme_minimal() 
  
  return(p)
}


# Loop through analytes per site to plot and save


chem_names <- unique(curry_cq$analyte)
for(analyte in chem_names){
  if(!is.na(analyte)){
   p <- plot_c_q_norm(curry_cq, analyte)
   #ggsave(p, path = "plots/cq", file=paste("Curry_", analyte, ".jpeg"), width = 20, height = 10, units = "in")
   print(p)
  }
}

chem_names <- unique(silv_cq$analyte)
for(analyte in chem_names){
  if(!is.na(analyte)){
    p <- plot_c_q_norm(silv_cq, analyte)
    #ggsave(p, file=paste("Silverado_", analyte, ".jpeg"), width = 20, height = 10, units = "in")
    print(p)
  }
}

chem_names <- unique(lakem_cq$analyte)
for(analyte in chem_names){
  if(!is.na(analyte)){
    p <- plot_c_q_norm(lakem_cq, analyte)
    #ggsave(p, file=paste("LakeM_", analyte, ".jpeg"), width = 20, height = 10, units = "in")
    print(p)
  }
}

```

## Precipitation

Precipitation data was collected from https://azmet.arizona.edu/azmet/27.htm - precipitation gage at the Desert Ridge station

```{r precip upload, warning=FALSE, message=FALSE, echo=FALSE}

precip <- read_csv(here("Data", "q_storms_precip.csv"))
precip <- precip %>% mutate(storm_date = as.Date(datetime)) %>% select(-`...1`)

hysteresis_df$storm_date <- as.Date(hysteresis_df$storm_date)

HI_precip <- left_join(hysteresis_df, precip, by = c("storm_date"))

curry_cq <- curry_cq %>%
  group_by(stormID) %>%
  mutate(
    storm_size = sum(precip_mm, na.rm = TRUE),
    storm_duration_hr = as.numeric(difftime(max(datetime, na.rm = TRUE), min(datetime, na.rm = TRUE), units = "hours")),
    storm_intensity = storm_size / storm_duration_hr
  ) %>%
  ungroup()
silv_cq <- silv_cq %>%
  group_by(stormID) %>%
  mutate(
    storm_size = sum(precip_mm, na.rm = TRUE),
    storm_duration_hr = as.numeric(difftime(max(datetime, na.rm = TRUE), min(datetime, na.rm = TRUE), units = "hours")),
    storm_intensity = storm_size / storm_duration_hr
  ) %>%
  ungroup()
lakem_cq <-lakem_cq %>%
  group_by(stormID) %>%
  mutate(
    storm_size = sum(precip_mm, na.rm = TRUE),
    storm_duration_hr = as.numeric(difftime(max(datetime, na.rm = TRUE), min(datetime, na.rm = TRUE), units = "hours")),
    storm_intensity = storm_size / storm_duration_hr
  ) %>%
  ungroup()

```


Time series of precip

```{r precip ts, warning=FALSE, message=FALSE, echo=FALSE}
precip <- precip %>% mutate(season = ifelse((month(datetime) >= 10 | month(datetime) < 5),  "winter", "summer"))
precip %>% filter(precip_mm < 700, precip_mm > 0) %>% ggplot(aes(x = datetime, y = precip_mm)) +
  geom_col(aes(color = season)) +
  theme_bw() +labs(title = "Daily Precip (excluding large outliers)")+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15), 
        plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 10))

precip %>% filter(precip_mm > 0)%>% ggplot(aes(x = datetime, y = precip_mm)) +
   geom_col(aes(color = season))+theme_bw() +labs(title = "Daily Precip all")+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15), 
        plot.title = element_text(size = 20), 
        axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 10))



```


## Scatterplot matricies


```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=10, out.height="100%"}
#need precip metrics from _cq df

curry_hi_beta_precip <- unique(curry_hi_beta %>% left_join(curry_cq %>% select(c(stormID, storm_size, storm_intensity, storm_duration_hr ,precip_7d_mm, precip_30d_mm, precip_90d_mm)), by ="stormID"))
silv_hi_beta_precip <- unique(silv_hi_beta %>% left_join(silv_cq %>% select(c(stormID, storm_size, storm_intensity, storm_duration_hr ,precip_7d_mm, precip_30d_mm, precip_90d_mm)), by ="stormID"))
lakem_hi_beta_precip <- unique(lakem_hi_beta %>% left_join(lakem_cq %>% select(c(stormID, storm_size, storm_intensity, storm_duration_hr ,precip_7d_mm, precip_30d_mm, precip_90d_mm)), by ="stormID"))

hi_beta_precip_all <- rbind(curry_hi_beta_precip, silv_hi_beta_precip, lakem_hi_beta_precip)
hi_beta_precip_all <- hi_beta_precip_all %>% filter(precip_30d_mm<1000, precip_90d_mm<1000, storm_size<1000, storm_intensity<100)#removing obvious outliers


# pivot wide so each analyte gets its own HI_mean/HI_sd/beta columns
storm_chars <- hi_beta_precip_all %>%
  select(site, stormID, analyte, HI_mean, HI_sd, beta,
         normalized_slope, max_width, peak_Q,
         precip_7d_mm, precip_30d_mm, precip_90d_mm, storm_size, storm_intensity) %>%
  pivot_wider(
    names_from = analyte,
    values_from = c(HI_mean, HI_sd, beta),
    names_glue = "{.value}_{analyte}"
  )

# build analyte-major column order
analyte_order <- unique(hi_beta_precip_all$analyte)

ordered_cols <- as.vector(sapply(analyte_order, function(a) {
  c(paste0("HI_mean_", a), paste0("HI_sd_", a), paste0("beta_", a))
}))

index_cols <- storm_chars %>% select(all_of(ordered_cols))

covariate_cols <- storm_chars %>%
  select(normalized_slope, max_width, peak_Q,
         precip_7d_mm, precip_30d_mm, precip_90d_mm, storm_size, storm_intensity)

covariate_labels <- c("normalized\nslope", "max\nwidth", "peak Q",
                       "precip\n7d", "precip\n30d", "precip\n90d", "storm\nsize", "storm\nintensity")

## splitting into 2 matricies for easier readability - beta and HI

hi_cols <- as.vector(sapply(analyte_order, function(a) {
  c(paste0("HI_mean_", a), paste0("HI_sd_", a))
}))

hi_index_cols <- storm_chars %>% select(all_of(hi_cols))

hi_cors <- cor(hi_index_cols, covariate_cols,
               method = "spearman",
               use = "pairwise.complete.obs")

colnames(hi_cors) <- covariate_labels
rownames(hi_cors) <- as.vector(sapply(analyte_order, function(a) {
  c(paste0("HI-", a), paste0("HI-", a, " (SD)"))
}))
fig_w <- 7
fig_h_hi <- fig_w * nrow(hi_cors) / ncol(hi_cors)


beta_cols <- paste0("beta_", analyte_order)

beta_index_cols <- storm_chars %>% select(all_of(beta_cols))

beta_cors <- cor(beta_index_cols, covariate_cols,
                  method = "spearman",
                  use = "pairwise.complete.obs")

colnames(beta_cors) <- covariate_labels
rownames(beta_cors) <- paste0("beta-", analyte_order)



```

\newpage

```{r, warning=FALSE, message=FALSE, echo=FALSE, out.width="50%", out.height="90%"}
corrplot(hi_cors,
         method = "ellipse",
         addCoef.col = "black",
         tl.col = "black",
         tl.cex = 0.5,        # axis label size
         number.cex =0.5,    # coefficient text size
         cl.pos = 'n',
         is.corr = FALSE,
         addrect = 16,
         rect.col = "black",
         rect.lwd = 5)

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=10, out.height="100%"}

corrplot(beta_cors,
         method = "ellipse",
         addCoef.col = "black",
         tl.col = "black",
         cl.pos = 'n',
         is.corr = FALSE,
         rect.lwd = 5)

```


### Individual scatterplots

```{r, fig.width=5, fig.height=4, fig.align="center", warning=FALSE, message=FALSE, echo=FALSE}

precip_vars <- c("precip_7d_mm", "precip_30d_mm", "precip_90d_mm", "storm_size", "storm_intensity")
metric_prefixes <- c("HI_mean", "HI_sd", "beta")

for (a in analyte_order) {
  for (metric in metric_prefixes) {
    
    y_col <- paste0(metric, "_", a)
    
    # skip if this analyte/metric combo doesn't exist as a column
    if (!y_col %in% names(storm_chars)) next
    
    for (p in precip_vars) {
      
      p_data <- storm_chars %>% select(all_of(y_col), all_of(p)) %>% na.omit()
      
      # skip if too few points to bother plotting
      if (nrow(p_data) < 3) next
      
      plt <- ggplot(p_data, aes(x = .data[[p]], y = .data[[y_col]])) +
        geom_point() +
        geom_smooth(method = "lm", se = FALSE, color = "blue") +
        labs(title = paste(y_col, "vs", p),
             x = p, y = y_col) +
        theme_minimal()
      
      print(plt)
    }
  }
}
```


\newpage 

## Precip and storm metrics

Storm size = cumulative precipitation (mm) per storm


```{r storm size v HI, warning=FALSE, message=FALSE, echo=FALSE}
curry_hi_beta_precip %>% filter(site == "Curry", analyte != "NA") %>% ggplot(aes(x = storm_size, y = HI_mean, color = season)) +
  geom_point() +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.05) +
  facet_wrap(~analyte) + labs(title = "Curry HI mean vs cumulative precip per storm", 
                              subtitle = "Error bars are 95% confidence intervals", x = "Storm size") +
  theme_bw()

silv_hi_beta_precip %>% filter(site == "Silverado", , analyte != "NA") %>% ggplot(aes(x = storm_size, y = HI_mean, color = season)) +
  geom_point() +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.05) +
  facet_wrap(~analyte) + labs(title = "Silverado HI mean vs cumulative precip per storm", 
                               subtitle = "Error bars are 95% confidence intervals", x = "Storm size") +
  theme_bw()


lakem_hi_beta_precip %>% filter(site == "Lake Margarita", , analyte != "NA") %>% ggplot(aes(x = storm_size, y = HI_mean, color = season)) +
  geom_point() +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.05) +
  facet_wrap(~analyte) + labs(title = "Lake Margarita HI mean vs cumulative precip per storm", 
                               subtitle = "Error bars are 95% confidence intervals", x = "Storm size") +
  theme_bw()
```

\newpage 

Storm intensity = cumulative precipitation/storm duration

```{r storm intensity v HI, warning=FALSE, message=FALSE, echo=FALSE}

curry_hi_beta_precip %>% filter(site == "Curry", analyte != "NA") %>% ggplot(aes(x = storm_intensity, y = HI_mean, color = season)) +
  geom_point() +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 0.05, color = "gray60", alpha = 0.05) +
  facet_wrap(~analyte) + labs(title = "Curry HI mean vs storm intensity", 
                              subtitle = "Error bars are 95% confidence intervals", x = "Storm Intensity") +
  theme_bw()

silv_hi_beta_precip %>% filter(site == "Silverado", , analyte != "NA") %>% ggplot(aes(x = storm_intensity, y = HI_mean, color = season)) +
  geom_point() +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 0.05, color = "gray60", alpha = 0.05) +
  facet_wrap(~analyte) + labs(title = "Silverado HI mean vs storm intensity", 
                               subtitle = "Error bars are 95% confidence intervals", x = "Storm Intensity") +
  theme_bw()


lakem_hi_beta_precip %>% filter(site == "Lake Margarita", , analyte != "NA") %>% ggplot(aes(x = storm_intensity, y = HI_mean, color = season)) +
  geom_point() +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 0.05, color = "gray60", alpha = 0.05) +
  facet_wrap(~analyte) + labs(title = "Lake Margarita HI mean vs storm intensity", 
                               subtitle = "Error bars are 95% confidence intervals", x = "Storm Intensity") +
  theme_bw()

```


