---
title: "Hysteresis Indicies and Beta of IBW sites: Curry, Silverado, and Lake Margarita"
output:
  pdf_document: default
  html_document:
    df_print: paged
indent: true
date: "2026-03-26"
---

Aims to plot hysteresis per analyte per storm per site

## Metrics



* *Hysteresis index* is calculated as the median (+/- 95% CI) across 2% intervals of discharge.


    +HI = clockwise (source limitation or nearstream source)


    -HI = counterclockwise (transport limitation or distal source)



* *Beta* is the slope of the CQ relationship on the rising limb of the storm

    CIs on beta are bootstrapped 95% CI

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=12, fig.height=8) 
```


```{r libraries, include=FALSE}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(here)
library(purrr)
```

## Data munging/cleaning

Workflow before this rmd, in order:

* q_storms.R: import discharge from Lake Margherite, Silverado, & Curry, delineate unique storms, and interpolate Q
* merge_Q_chems.R: merge discharge data with chem data per site, produce brief cq plots per analyte per site
* hysteresis_calc_2.R: calculate hysteresis based on python function, removed a lot of storms with not enough data

```{r load data, include=FALSE}

# note - we cannot access google drive API when knitting rmd, so these files will have to be downloaded locally. this can be done by running hysteresis_calc_2.R

curry_hysteresis <- readRDS(here("Data/curry_hysteresis.rds")) # raw product from hysteresis calculation, per analyte per storm
silv_hysteresis <- readRDS(here("Data/silv_hysteresis.rds"))
lakem_hysteresis <- readRDS(here("Data/lakem_hysteresis.rds"))

hysteresis_df <- read.csv(here("Data/hysteresis_summary.csv")) # summearized, combined all sites, includes mean HI per analyte per storm, amongst other things
hi_intervals_df <- read.csv(here("Data/hysteresis_intervals.csv")) # raw HI per interval per site per storm per analyte
beta_df <- read.csv(here("Data/beta_all_sites.csv"))#beta per storm per analyte per site

# from google drive - no API access needed 
curry_cq <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", 
                             "1wWUPyysrln9gyWpfS6vAt2u-MS-LtOyT")) %>% 
  select(-X) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"))
curry_cq[1,1] <- "2008-01-29 00:00:00"

silv_cq <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", 
                             "1J-mW-MfwTfq7sOsy2sn4eejb9cq0ykvR")) %>% 
  select(-X) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"))
silv_cq[1,1] <- "2008-01-29 00:00:00"

lakem_cq <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", 
                            "100HPPcicy6tv7CUrwzDI8Yh4C6FBozhG")) %>% 
  select(-X) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"))
lakem_cq[1,1] <- "2008-01-29 00:00:00"
```

```{r data munging, message=FALSE, warning=FALSE, include=FALSE}
#some quick data munging - need to get storm_date into beta
storm_date <- hysteresis_df %>% select(c(site, stormID, storm_date, analyte))
beta_df <- left_join(beta_df, storm_date, by = c("site", "stormID", "analyte"))
```


## Plotting Hysteresis

```{r echo=FALSE, warning=FALSE, message=FALSE}
hi_summary <- hysteresis_df %>%
  group_by(site, analyte) %>%
  summarise(
    median_HI = median(HI_mean, na.rm = TRUE),
    lower_CI  = quantile(HI_mean, 0.25, na.rm = TRUE),  # 25th percentile
    upper_CI  = quantile(HI_mean, 0.75, na.rm = TRUE),  # 75th percentile
    n         = n(),
    .groups = "drop"
  )
hi_summary <- hi_summary %>% filter(!is.na(analyte))
hysteresis_df %>% filter(!is.na(analyte)) %>%  ggplot() +
  geom_jitter(aes(x = site, y = HI_mean),
              width = 0.15, alpha = 0.3, size = 1.5) +
  geom_pointrange(data = hi_summary,
                  aes(x = site, y = median_HI, ymin = lower_CI, ymax = upper_CI),
                  size = 0.8, linewidth = 1.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Median HI by Analyte and Site",
       subtitle = "Points = individual storms, range = IQR",
       x = NULL, y = "HI") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))


#per site
hysteresis_df %>%
  filter(site == "Curry", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Curry - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

hysteresis_df %>%
  filter(site == "Silverado", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Silverado - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

hysteresis_df %>%
  filter(site == "Lake Margarita", !is.na(analyte)) %>%
  ggplot(aes(x = as.Date(storm_date), y = HI_mean)) +
  geom_errorbar(aes(ymin = HI_mean - HI_sd, ymax = HI_mean + HI_sd), 
                width = 10, color = "gray60", alpha = 0.7) +
  geom_point() +
  #geom_smooth(method = "loess", se = FALSE, color = "red", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ analyte, scales = "free_y") +
  labs(title = "Lake Margarita - mean HI with confidence intervals per storm per analyte",
       x = "Storm Date", y = "Mean HI", size = "Peak Q") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Plotting Beta
```{r plotting beta, echo=FALSE, warning=FALSE, message=FALSE}

beta_df %>% filter(!is.na(analyte), analyte != "Pb", outlier==F) %>% ggplot(aes(x = as.Date(storm_date), y = beta, color = site))+
  geom_point()+
  geom_errorbar(aes(ymin = boot_ci_lo, ymax = boot_ci_hi),
                width = 10, alpha = 0.4) +
  facet_wrap(~analyte, scales = "free_y") + theme_minimal() +
  labs(title = "Beta per storm per analyte per site", x = "Storm Date", subtitle = "Bars are 95% confidence intervals")

```



## Plotting CQ
```{r cq plots, echo=FALSE, message=FALSE, warning=FALSE}
## attempting to make function for this - from sycburns
plot_c_q_norm <- function(cq_data, analyte_col) {
  curr_site <- cq_data %>% 
    filter(analyte == analyte_col) %>% 
    filter( !is.na(cfs)) 
  
  if (nrow(curr_site) == 0) {
    message(paste("Skipping", site, "- no valid data"))
    return(NULL)
  }
  
  if (!"Time_hours" %in% names(curr_site)) {
    curr_site <- curr_site %>%
     # group_by(stormID) %>% 
      mutate(Time_hours = as.numeric(difftime(datetime, min(datetime), units = "hours")))
  }
  # # Remove top 10% of cfs values per Event_Date (storm)
  # curr_site <- curr_site %>%
  #   group_by(Event_Date) %>%
  #   filter(cfs <= quantile(cfs, 0.9, na.rm = TRUE)) %>%
  #   ungroup()
  
  #Normalize within each storm
  curr_site <- curr_site %>%
    group_by(stormID) %>%
    mutate(
      Level_norm = (cfs - min(cfs, na.rm = TRUE)) /
        (max(cfs, na.rm = TRUE) - min(cfs, na.rm = TRUE)),
      Chem_norm = (.data$mean_conc - min(.data$mean_conc, na.rm = TRUE)) /
        (max(.data$mean_conc, na.rm = TRUE) - min(.data$mean_conc, na.rm = TRUE)),
      Time_norm = (Time_hours - min(Time_hours, na.rm = TRUE)) /
        (max(Time_hours, na.rm = TRUE) - min(Time_hours, na.rm = TRUE))
    ) %>%
    ungroup() %>%
    filter(!is.na(.data$mean_conc), !is.na(cfs)) 
  
  #Plot normalized cq relationship per storm
  p <- curr_site %>% filter(!is.na(curr_site$Level_norm), !is.na(curr_site$Chem_norm)) %>% ggplot(aes(x = Level_norm, y = Chem_norm)) + 
    geom_point(aes(color = Time_norm)) + 
    facet_wrap(~stormID, scales = "free") + 
    scale_color_gradientn(colours = rainbow(7)) + 
    labs( title = paste(curr_site$Site[1], "-", analyte_col), 
          x = "Normalized Stage (0–1)", 
          y = paste("Normalized", analyte_col, "(0–1)"), 
          color = "Time (within storm)" ) + 
    theme_minimal() 
  
  return(p)
}


# Loop through analytes per site to plot and save
setwd(here("plots", "cq"))

chem_names <- unique(curry_cq$analyte)
for(analyte in chem_names){
  if(!is.na(analyte)){
   p <- plot_c_q_norm(curry_cq, analyte)
   #ggsave(p, file=paste("Curry_", analyte, ".jpeg"), width = 20, height = 10, units = "in")
   print(p)
  }
}

chem_names <- unique(silv_cq$analyte)
for(analyte in chem_names){
  if(!is.na(analyte)){
    p <- plot_c_q_norm(silv_cq, analyte)
    #ggsave(p, file=paste("Silverado_", analyte, ".jpeg"), width = 20, height = 10, units = "in")
    print(p)
  }
}

chem_names <- unique(lakem_cq$analyte)
for(analyte in chem_names){
  if(!is.na(analyte)){
    p <- plot_c_q_norm(lakem_cq, analyte)
    #ggsave(p, file=paste("LakeM_", analyte, ".jpeg"), width = 20, height = 10, units = "in")
    print(p)
  }
}

```


